---
title: "LAB11"
author: "Bryce Butler"
date: "2024-11-08"
output: 
  html_document:
    toc: yes
    toc_float: yes
    css: "custom.css"
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FALL224753butl0072)
```


# Task 1 
```{r}
getwd()
```

# Task 2 

The \( 95\% \) confidence interval for the mean \( \mu \) is:

\[
\bar{y} \pm t_{\frac{\alpha}{2}} \frac{s}{\sqrt{n}}
\]



The confidence interval for the population variance \( \sigma^2 \) is given by:

\[
\left( \frac{(n - 1)s^2}{\chi^2_{\frac{\alpha}{2}}}, \frac{(n - 1)s^2}{\chi^2_{1 - \frac{\alpha}{2}}} \right)
\]


```{r}
# Sample data
d <- c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565,
       4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781,
       5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185,
       4.9879)

# Confidence Intervals for the Mean
mean_ci_95 <- t.test(d, conf.level = 0.95)$conf.int
mean_ci_90 <- t.test(d, conf.level = 0.90)$conf.int
mean_ci_80 <- t.test(d, conf.level = 0.80)$conf.int
mean_ci_50 <- t.test(d, conf.level = 0.50)$conf.int

# 80% CI using t.test and extracting values
t_test_80 <- t.test(d, conf.level = 0.80)
mean_ci_80_manual <- t_test_80$conf.int

# Confidence Intervals for the Variance
n <- length(d)
s2 <- var(d)

# Function to calculate variance CI
variance_ci <- function(alpha) {
  chi2_lower <- qchisq(alpha / 2, df = n - 1, lower.tail = FALSE)
  chi2_upper <- qchisq(1 - alpha / 2, df = n - 1, lower.tail = FALSE)
  lower <- ((n - 1) * s2) / chi2_lower
  upper <- ((n - 1) * s2) / chi2_upper
  return(c(lower, upper))
}

# Variance Confidence Intervals
var_ci_95 <- variance_ci(0.05)
var_ci_90 <- variance_ci(0.10)
var_ci_80 <- variance_ci(0.20)
var_ci_50 <- variance_ci(0.50)

# Displaying Results
cat("95% CI for the Mean:", mean_ci_95, "\n")
cat("90% CI for the Mean:", mean_ci_90, "\n")
cat("80% CI for the Mean:", mean_ci_80, "\n")
cat("50% CI for the Mean:", mean_ci_50, "\n")

cat("95% CI for the Variance:", var_ci_95, "\n")
cat("90% CI for the Variance:", var_ci_90, "\n")
cat("80% CI for the Variance:", var_ci_80, "\n")
cat("50% CI for the Variance:", var_ci_50, "\n")
```

# Task 3 

For large independent samples with \( n_1, n_2 \geq 30 \), the confidence interval for the difference in means \( \mu_1 - \mu_2 \) is:

\[
\bar{y}_1 - \bar{y}_2 \pm z_{\frac{\alpha}{2}} \sqrt{\frac{\sigma_1^2}{n_1} + \frac{\sigma_2^2}{n_2}}
\]

If we assume equal variances, the confidence interval becomes:

\[
\bar{y}_1 - \bar{y}_2 \pm z_{\frac{\alpha}{2}} \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}
\]

For small sample sizes \( n_1 \) or \( n_2 < 30 \), and assuming equal variances \( \sigma_1 = \sigma_2 \), the confidence interval is:

\[
\bar{y}_1 - \bar{y}_2 \pm t_{\frac{\alpha}{2}, \, df} \sqrt{s_p^2 \left( \frac{1}{n_1} + \frac{1}{n_2} \right)}
\]

where \( s_p^2 \) is the pooled variance:

\[
s_p^2 = \frac{(n_1 - 1)s_1^2 + (n_2 - 1)s_2^2}{n_1 + n_2 - 2}
\]

```{r}
# Sample data
snapper <- c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93,
             25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

blue <- c(21.65, 17.48, 20.11, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93,
          15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 16.5, 19.03)

# 95% Confidence Interval for difference in means (assuming equal variances)
t_test_equal_var <- t.test(snapper, blue, var.equal = TRUE, conf.level = 0.95)
ci_95 <- t_test_equal_var$conf.int

# Practical interpretation
cat("The 95% confidence interval for the difference in means (µ_snapper - µ_blue) is:",
    round(ci_95[1], 4), "to", round(ci_95[2], 4), "\n")

# Additional confidence intervals for different levels
ci_85 <- t.test(snapper, blue, var.equal = TRUE, conf.level = 0.85)$conf.int
ci_75 <- t.test(snapper, blue, var.equal = TRUE, conf.level = 0.75)$conf.int
ci_25 <- t.test(snapper, blue, var.equal = TRUE, conf.level = 0.25)$conf.int

# Displaying Results
cat("85% CI for (µ_snapper - µ_blue):", round(ci_85[1], 4), "to", round(ci_85[2], 4), "\n")
cat("75% CI for (µ_snapper - µ_blue):", round(ci_75[1], 4), "to", round(ci_75[2], 4), "\n")
cat("25% CI for (µ_snapper - µ_blue):", round(ci_25[1], 4), "to", round(ci_25[2], 4), "\n")

# Observing the effect of confidence level on interval width
cat("As the confidence level decreases, the width of the confidence interval also decreases.\n")
```


# Task 4 


For paired samples (dependent samples) with a large sample size \( n \geq 30 \), the confidence interval for the difference in means \( \mu_d = (\mu_1 - \mu_2) \) is:

\[
d \pm z_{\frac{\alpha}{2}} \frac{\sigma_d}{\sqrt{n}}
\]

For smaller sample sizes \( n < 30 \), the confidence interval is:

\[
d \pm t_{\frac{\alpha}{2}, \, n-1} \frac{s_d}{\sqrt{n}}
\]

where \( s_d \) is the sample standard deviation of the differences.


```{r}
# Sample data
Exam1 <- c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89,
           78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65,
           61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.60)

Exam2 <- c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55,
           87.18, 63.82, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44,
           74.37, 75.82, 61.86, 59.92, 64.42, 73.57, 57.15)

# Paired t-test for the difference (Exam2 - Exam1) with 95% CI
t_test_95 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.95)
ci_95 <- t_test_95$conf.int

# Practical interpretation
cat("The 95% confidence interval for the difference in means (µ_d = µ_Exam2 - µ_Exam1) is:",
    round(ci_95[1], 4), "to", round(ci_95[2], 4), "\n")

# Additional confidence intervals for different confidence levels
ci_90 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.90)$conf.int
ci_80 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.80)$conf.int
ci_70 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.70)$conf.int
ci_60 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.60)$conf.int
ci_10 <- t.test(Exam2, Exam1, paired = TRUE, conf.level = 0.10)$conf.int

# Displaying Results
cat("90% CI for µ_d:", round(ci_90[1], 4), "to", round(ci_90[2], 4), "\n")
cat("80% CI for µ_d:", round(ci_80[1], 4), "to", round(ci_80[2], 4), "\n")
cat("70% CI for µ_d:", round(ci_70[1], 4), "to", round(ci_70[2], 4), "\n")
cat("60% CI for µ_d:", round(ci_60[1], 4), "to", round(ci_60[2], 4), "\n")
cat("10% CI for µ_d:", round(ci_10[1], 4), "to", round(ci_10[2], 4), "\n")

# Observing the effect of confidence level on interval width
cat("As the confidence level decreases, the width of the confidence interval also decreases.\n")
```

# Task 5

For a single sample with \( n \geq 30 \), the confidence interval for a population proportion \( p \) is:

\[
\hat{p} \pm z_{\frac{\alpha}{2}} \sqrt{\frac{\hat{p}(1 - \hat{p})}{n}}
\]

For two independent samples with large sample sizes \( n_1, n_2 \geq 30 \), the confidence interval for the difference in proportions \( p_1 - p_2 \) is:

\[
\hat{p}_1 - \hat{p}_2 \pm z_{\frac{\alpha}{2}} \sqrt{\frac{\hat{p}_1 (1 - \hat{p}_1)}{n_1} + \frac{\hat{p}_2 (1 - \hat{p}_2)}{n_2}}
\]

where:
- \( \hat{p}_1 \) is the sample proportion from the first sample,
- \( \hat{p}_2 \) is the sample proportion from the second sample.

```{r}
# Sample data from the NZBIRDS dataset
extinct_sample_size <- 38
extinct_flightless <- 21

non_extinct_sample_size <- 78
non_extinct_flightless <- 7

# Proportions of flightless birds in each group
p_hat_extinct <- extinct_flightless / extinct_sample_size
p_hat_non_extinct <- non_extinct_flightless / non_extinct_sample_size

# Difference in proportions
prop_diff <- p_hat_extinct - p_hat_non_extinct

# Standard error for the difference in proportions
se_diff <- sqrt((p_hat_extinct * (1 - p_hat_extinct) / extinct_sample_size) +
                (p_hat_non_extinct * (1 - p_hat_non_extinct) / non_extinct_sample_size))

# Z-score for 95% confidence level
z_score <- qnorm(0.975)

# 95% confidence interval for the difference in proportions
ci_lower <- prop_diff - z_score * se_diff
ci_upper <- prop_diff + z_score * se_diff

# Displaying the result
cat("The 95% confidence interval for the difference in proportions is:",
    round(ci_lower, 4), "to", round(ci_upper, 4), "\n")
```

# Task 6


The confidence interval for the ratio of two population variances, assuming two independent samples from normal populations, is given by:

\[
\frac{s_1^2}{s_2^2} \cdot \frac{1}{F_{\frac{\alpha}{2}, \, \nu_1, \, \nu_2}} < \frac{\sigma_1^2}{\sigma_2^2} < \frac{s_1^2}{s_2^2} \cdot F_{\frac{\alpha}{2}, \, \nu_1, \, \nu_2}
\]

where:
- \( s_1^2 \) and \( s_2^2 \) are the sample variances,
- \( F_{\frac{\alpha}{2}, \, \nu_1, \, \nu_2} \) is the critical value from the F-distribution with degrees of freedom \( \nu_1 = n_1 - 1 \) and \( \nu_2 = n_2 - 1 \),
- \( \alpha \) is the significance level, and \( 1 - \alpha \) is the confidence level.


For the calculated confidence interval, interpret the results:

95% confident that the true ratio of variances between the two populations lies within the calculated interval

```{r}
# Simulated data
set.seed(35)
sam1 <- rnorm(25, mean = -10, sd = 5)

set.seed(45)
sam2 <- rnorm(34, mean = 40, sd = 8)

# 95% Confidence Interval for the ratio of variances
var_test_95 <- var.test(sam1, sam2, conf.level = 0.95)
ci_95 <- var_test_95$conf.int

cat("The 95% confidence interval for the ratio of variances (σ1^2 / σ2^2) is:",
    round(ci_95[1], 4), "to", round(ci_95[2], 4), "\n")

# Additional confidence intervals at different confidence levels
ci_80 <- var.test(sam1, sam2, conf.level = 0.80)$conf.int
ci_70 <- var.test(sam1, sam2, conf.level = 0.70)$conf.int
ci_60 <- var.test(sam1, sam2, conf.level = 0.60)$conf.int
ci_50 <- var.test(sam1, sam2, conf.level = 0.50)$conf.int

# Displaying Results
cat("80% CI for the ratio of variances (σ1^2 / σ2^2):", round(ci_80[1], 4), "to", round(ci_80[2], 4), "\n")
cat("70% CI for the ratio of variances (σ1^2 / σ2^2):", round(ci_70[1], 4), "to", round(ci_70[2], 4), "\n")
cat("60% CI for the ratio of variances (σ1^2 / σ2^2):", round(ci_60[1], 4), "to", round(ci_60[2], 4), "\n")
cat("50% CI for the ratio of variances (σ1^2 / σ2^2):", round(ci_50[1], 4), "to", round(ci_50[2], 4), "\n")

# Observing the effect of confidence level on interval width
cat("As the confidence level decreases, the width of the confidence interval for the ratio of variances also decreases.\n")
```


# Task 7
```{r}
# Set the seed and generate a sample
set.seed(23)
x <- rnorm(30, mean = 10, sd = 12)

# Call the myci function from your package
FALL224753butl0072::myci(x)
```
