---
title: "LAB9"
author: "Bryce Butler"
date: "2024-10-29"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FALL224753butl0072)
load("~/Downloads/PROJECTS/FALL224753butl0072/data/spruce.df.rda")
```

# Task 1 
```{r}
getwd()
```

# Task 2 

WHY SHOULD EACH DATUM IN X BE SELECTED WITH EQUAL OPPORTUNITY? 

It is necessary to ensure that each data point has an equal chance of being selected for each bootstrap sample
```{r}
# Sampling test
set.seed(35)
sam <- round(rnorm(20, mean=10, sd=4), 2)

# With replacement
for (i in 1:5) {
  print(unique(sample(sam, 20, replace=TRUE)))
}

# Without replacement
for (i in 1:5) {
  print(unique(sample(sam, 20, replace=FALSE)))
}

# Attempting to sample more than available elements without replacement
sample(sam, 21, replace=TRUE)  
```
An error is thrown when replace is set to FALSE because it is attempting to sample more than is available without replacement. 


# Task 3

In each of the above cases how close is the point estimate to the population value?

The point estimate is very close to the population value. 

```{r}
# Define the `myboot2` function
myboot2 <- function(iter=10000, x, fun="mean", alpha=0.05, cx=1.5, ...) {  
  n <- length(x)
  y <- sample(x, n * iter, replace=TRUE)
  rs.mat <- matrix(y, nr=n, nc=iter, byrow=TRUE)
  xstat <- apply(rs.mat, 2, fun)
  ci <- quantile(xstat, c(alpha/2, 1-alpha/2))
  
  para <- hist(xstat, freq=FALSE, las=1,
               main=paste("Histogram of Bootstrap Sample Statistics", "\n",
                          "alpha=", alpha, " iter=", iter, sep=""), 
               ...)
  mat <- matrix(x, nr=length(x), nc=1, byrow=TRUE)
  pte <- apply(mat, 2, fun)
  
  abline(v=pte, lwd=3, col="Black")
  segments(ci[1], 0, ci[2], 0, lwd=4)
  text(ci[1], 0, paste("(", round(ci[1], 2), sep=""), col="Red", cex=cx)
  text(ci[2], 0, paste(round(ci[2], 2), ")", sep=""), col="Red", cex=cx)
  text(pte, max(para$density) / 2, round(pte, 2), cex=cx)
  
  invisible(list(ci=ci, fun=fun, x=x))
}

# Set seeds and generate samples for mean estimation
set.seed(39); sam1 <- rnorm(25, mean=25, sd=10)
set.seed(30); sam2 <- rchisq(20, df=3)
set.seed(40); sam3 <- rgamma(30, shape=2, scale=3)
set.seed(10); sam4 <- rbeta(20, shape1=3, shape2=4)

# 95% CI for the mean
myboot2(iter=10000, x=sam1, fun="mean", alpha=0.05)
myboot2(iter=10000, x=sam2, fun="mean", alpha=0.05)
myboot2(iter=10000, x=sam3, fun="mean", alpha=0.05)
myboot2(iter=10000, x=sam4, fun="mean", alpha=0.05)

# 80% CI for the variance
myboot2(iter=10000, x=sam1, fun="var", alpha=0.20)
myboot2(iter=10000, x=sam2, fun="var", alpha=0.20)
myboot2(iter=10000, x=sam3, fun="var", alpha=0.20)
myboot2(iter=10000, x=sam4, fun="var", alpha=0.20)
```

# Task 4 
```{r}
# Update `myboot2` to return `xstat`
myboot2 <- function(iter=10000, x, fun="mean", alpha=0.05, cx=1.5, ...) {  
  n <- length(x)
  y <- sample(x, n * iter, replace=TRUE)
  rs.mat <- matrix(y, nr=n, nc=iter, byrow=TRUE)
  xstat <- apply(rs.mat, 2, fun)
  ci <- quantile(xstat, c(alpha/2, 1-alpha/2))
  
  para <- hist(xstat, freq=FALSE, las=1,
               main=paste("Histogram of Bootstrap Sample Statistics", "\n", 
                          "alpha=", alpha, " iter=", iter, sep=""), 
               ...)
  mat <- matrix(x, nr=length(x), nc=1, byrow=TRUE)
  pte <- apply(mat, 2, fun)
  
  abline(v=pte, lwd=3, col="Black")
  segments(ci[1], 0, ci[2], 0, lwd=4)
  text(ci[1], 0, paste("(", round(ci[1], 2), sep=""), col="Red", cex=cx)
  text(ci[2], 0, paste(round(ci[2], 2), ")", sep=""), col="Red", cex=cx)
  text(pte, max(para$density) / 2, round(pte, 2), cex=cx)
  
  invisible(list(ci=ci, fun=fun, x=x, xstat=xstat))
}

# Calculate median for the provided sample and plot xstat
sam <- c(1,1,1,2,2,2,2,3,3,3,4,4)
result <- myboot2(x=sam, fun="median")
barplot(result$xstat, main="Bootstrap Sample Statistics for Median")
result$ci
```

Boostrap Estimate for the mean (L, U): 1.5, 3.0


# Task 5
```{r}
mean_median_ratio <- function(sam, alpha=0.05) {
  myboot2(x=sam, fun=function(x) mean(x)/median(x), alpha=alpha)
}

# Samples
set.seed(39); sam1 <- rnorm(25, mean=25, sd=10)
set.seed(30); sam2 <- rchisq(20, df=3)
set.seed(40); sam3 <- rgamma(30, shape=2, scale=3)
set.seed(10); sam4 <- rbeta(20, shape1=3, shape2=4)

# 95% and 70% Intervals
mean_median_ratio(sam1, alpha=0.05)
mean_median_ratio(sam2, alpha=0.05)
mean_median_ratio(sam3, alpha=0.05)
mean_median_ratio(sam4, alpha=0.05)

mean_median_ratio(sam1, alpha=0.30)
mean_median_ratio(sam2, alpha=0.30)
mean_median_ratio(sam3, alpha=0.30)
mean_median_ratio(sam4, alpha=0.30)
```

# Task 6
```{r}
# Additional distributions
set.seed(50); sam5 <- rexp(20, rate=1)
set.seed(51); sam6 <- rgeom(20, prob=0.5)
set.seed(52); sam7 <- rlogis(20, location=0, scale=1)
set.seed(53); sam8 <- rweibull(20, shape=2, scale=1)

# 80% CI for mean and variance
myboot2(x=sam5, fun="mean", alpha=0.20)
myboot2(x=sam5, fun="var", alpha=0.20)
myboot2(x=sam6, fun="mean", alpha=0.20)
myboot2(x=sam6, fun="var", alpha=0.20)
myboot2(x=sam7, fun="mean", alpha=0.20)
myboot2(x=sam7, fun="var", alpha=0.20)
myboot2(x=sam8, fun="mean", alpha=0.20)
myboot2(x=sam8, fun="var", alpha=0.20)
```

# Task 7
```{r}
set.seed(68)
sam <- rnorm(20, mean=10, sd=4)

# Theoretical 95% CI
theoretical_ci <- c(mean(sam) - qnorm(0.975) * sd(sam) / sqrt(length(sam)), 
                    mean(sam) + qnorm(0.975) * sd(sam) / sqrt(length(sam)))
theoretical_ci

# Bootstrap 95% CI
myboot2(x=sam, fun="mean", alpha=0.05)
```

How do they compare? 

They are very close. They both contain the mean in the interval, but the bootstrap could be considered a tad bit more accurate. The bootsrap is more effective since it does not assume normally and has more flexibility. 


# Task 8
```{r}
obj <- FALL224753butl0072::myboot2(x =spruce.df$pH)
print(obj)
```
